### 2011-10-17 doc/manuscript_plos.tex
A PLoS template is added with the same content as one in manuscript.tex.

\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in}
\setlength{\parskip}{\bigskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[authoryear]{natbib}
\usepackage{longtable}
\doublespacing

\newcommand{\chck}{\textbf{[CHECK]}}
\newcommand{\addfig}{\textbf{[ADD FIGURE]}}
\newcommand{\tba}{\textbf{[TO BE ADDED]}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newcommand{\lyxaddress}[1]{
\par {\raggedright #1
\vspace{1.4em}
\noindent\par}
}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{babel}

\begin{document}
\renewcommand\refname{Literature Cited}

\raggedright

\subsubsection*{Title: }

Inferring homologous recombination and horizontal gene transfer 
using whole genomes of \textit{Streptococcus pyogenes} 
and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}

\subsubsection*{Authors:}

Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, Melissa Jane Hubisz$^{1}$, Ilan 
Gronau$^{1}$,
Michael J. Stanhope$^{2}$, and Adam Siepel$^{1}$

\subsubsection*{Affiliations:}

\lyxaddress{$^{1}$Department of Biological Statistics and Computational Biology,
Cornell University, Ithaca, NY 14853, USA.}

\lyxaddress{$^{2}$Department of Population Medicine and Diagnostic Sciences,
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853,
USA.}

\subsubsection*{Author for Correspondence:}

Adam Siepel, Department of Biological Statistics and Computational
Biology, Cornell University, Ithaca, NY 14853, USA, Phone: 607-254-1157,
Fax: 607-255-4698, Email: acs4@cornell.edu\clearpage{}

\subsubsection*{Abbreviations:}

group A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
\textit{equisimilis} (SDE)

\subsubsection*{Keywords:}

bacterial recombination, gene duplication and transfer, ClonalOrigin,
ClonalFrame, Mowgli

\subsubsection*{Version: 0.68 (Oct 15, 2011) \clearpage{}}

\section*{Abstract}

@@

\section{Introduction}

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}. However, it has been reported to cause serious and
life-threatening streptococcal diseases to humans traditionally associated with
GAS \citep[e.g.,][]{Brandt2009}.  Because of the overlap in disease profile and
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
well interact with each other by exchanging parts of their genomes.

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal of
genomes available.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond such rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of \citet{Karlin2001}. \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to \citet{Zhaxybayeva2006}.  \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
\citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

@@

Homologous recombining gene transfer could be studied more readily using the
core genomes of individual bacteria while horizontal gene transfer might be more
easily detected using the pan genomes. Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segments between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.  Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by \citet{Doyon2011}.  We noted that the first
approach could be used to infer transfer of genomic segments due to homologous
recombination, and the second could be useful in studying horizontal gene
transfer.

@@

\section{Results}

\subsection{Recombination drived more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
model-based approach.  A genome alignment of the five genomes showed two
distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
SDD genomes were larger in length than those of SPY. Although the graphical view
of the genome alignment did not clearly show whether SDD was different from
either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
the genome alignment showed that the external SDD branch was attached to the
internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
distribution of the alignment block lengths was skewed to the right with mean of
4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
model assumption of geometric block length distribution might be plausible in
\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
\citealt{Didelot2010}).  The relative strength ($r/m$) of recombination compared
to mutations indicated that recombination-driven substitutions affected more
sites of the core genomes than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
species bacterial recombination process appeared to accelerate molecular
evolution of more sites than did point mutations process.

\subsection{Recombination hotspots were observed in genomes}

@@

\subsection{SPY and SDE were more closely related in about tenth of the core genome}

@@

\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}

@@

\subsection{Reduction of genes and divergence of SPY and SDE species}

@@

\subsection{Virulence genes via recombining gene transfer}

@@

\subsection{Association of virulence genes and horizontal gene transfer}

@@

\subsection{Model-based approach or parsimony-based one for recombining gene
transfer study}

@@

\subsection{UCSC genome browser tracks for bacterial gene transfer study}

@@

\subsection{Units of recombining gene transfer}

@@

\subsection{Transporters, transferases, aminoacyl tRNA synthetases, and
transposases}

@@

\subsection{Simulations}

@@

\section{Discussions}

@@

Were there unbalanced recombining gene transfer between the two species, what
were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
the species?  \citet{Kalia2001} discussed restriction-modification system (RMS)
as a possible mechanism of the unbalanced gene flow.  Restriction-modification
system is a biological self-defense mechanism of bacteria, which is diverse in
prokaryote world.  Biological roles of RMS might be to maintain and control
species identity \citep{Jeltsch2003}.  Recently, \citet{Budroni2011a} studied
that \textit{Neisseria meningitidis} phylogenetic clades were associated with
RMS, claiming that RMS modulated homologous recombination in the bacteria of
\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contained
restriction-modification type I system as well as type II system. The two
genomes of SDE contained only type II system. More tight self-defense system of
SPY may have made the species less capable of aquiring foreign genetic material
from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
may have been associated with different restriction-modification systems in the
two species \citep{Kalia2001}.

We reexamined the seven genes that \citet{Kalia2001} had studied for net
directionality of gene transfer between SPY and SDE.  The internal fragments of
the six housekeeping genes (\textit{gki}, \textit{gtr}, \textit{murI},
\textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
we investigated in this study.  Except for \textit{mutS} where \citet{Kalia2001}
did not find recombination, our results were consistent with that of
\citet{Kalia2001}.  \citet{Pinho2010} reported recombination in the gene
\textit{parC}, in which we also observed moderate recombination.

@@

\section{Materials and Methods}

\subsection{Complete bacterial genomes}

Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
accessions: NC\_004070 \citep{Beres2002}, NC\_008024 \citep{Beres2006},
NC\_012891 \citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to
these samples as SPY1, SPY2, SDE1, and SDE2, respectively. The genome of
\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
\citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
ssp.\textit{\ equi} strain 4047 referred to as SEE \citep{Holden2009}.

\subsection{Model-based method for recombining gene transfer}

We present a short summary of the statistical method developed by
\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
recombination. A detailed account was found in the two papers and in the user
guide available at \url{http://code.google.com/p/clonalorigin}.  We used
\texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with default
options to attempt to align the six genomes, which led to too small proportion
of core genomes. We had to remove the SEE sample from the genome alignment
although we kept it in the parsimony-based approach.  We then used
\texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
identified 276 blocks of sequence alignments with length threshold of 1500 base
pairs.  We excluded two of these blocks in a subsequent analysis because they
contained long consecutive gaps.  We applied \texttt{ClonalFrame} v1.1 available
at \url{http://www.xavierdidelot.xtreemhost.com} to the remaining 274 blocks,
and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
also estimates the fraction between recombination and mutation rates.
We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
and 10000 sampling generations to obtain the tree estimate with branch lengths,
and additionally seven independent MCMC runs to validate adequate convergence of
the first run.

Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
employed independently for each of the 274 alignment blocks as described in
\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
recombination rate, and recombinant tract length.  Two independent chains were
executed with $10^6$ burn-in generations and $10^7$ sampling generations. We
sub-sampled the result every $10^5$-th generations.  A few alignment blocks were
not successfully finished even after a month of computation.  Because those were
not expected to affect the global estimates of population parameters
significantly, we used only finished blocks to estimate mutation rate,
recombination rate, and recombinant tract length.  The global estimators of the
three population parameters were medians weighted over block lengths, which
down-weighed less reliable estimates obtained from shorter alignment blocks.
One of the two independent chains was used to obtain the block-wise mean
estimates, and the other chain was used to validate convergence of the first
stage of MCMC.  Two independent MCMC runs resulted in similar estimates of the
three population parameters (Table S\ref{tab:three}).  The second stage of MCMC
with \texttt{ClonalOrigin} was executed with the three population parameters
fixed to the estimated values in the first stage.  Two independent MCMC chains
were also executed with a total generation being $1.1\times10^8$. The first
$10^7$ generations of each chain were discarded as burn-in. The remaining $10^8$
generations were sub-sampled every $10^5$-th generation.

\subsection{Parsimony-based method for horizontal gene transfer}

>>>>

% Four potential reviewers whose e-mail addresses you can provide
% Claus O. Wilke
% Debra Bessen
% Bruce R. Levin
% Michael Wessels

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace} 
\doublespacing

% FIXME: Remove this package when submitting the manuscript.
\usepackage{url}

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}

% FIXME: Remove the followings if you have to when submitting the manuscript.
\let\citep\cite
\let\citet\cite
\providecommand{\tabularnewline}{\\}
\newcommand{\lyxdot}{.}
\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}

\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Inferring homologous recombination and horizontal gene transfer 
using whole genomes of \textit{Streptococcus pyogenes} 
and\textit{ S.\ dysgalactiae} ssp.\textit{\ equisimilis}}
}
% Short Title (running head): Gene transfer using Streptococcus genomes
% Insert Author names, affiliations and corresponding author email.
\\
Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, 
Melissa Jane Hubisz$^{1}$, Ilan Gronau$^{1}$,
Michael J. Stanhope$^{2}$, and Adam Siepel$^{1,\ast}$
\\
\bf{1} 
Department of Biological Statistics and Computational Biology,
Cornell University, Ithaca, NY 14853, USA
\\
\bf{2}
Department of Population Medicine and Diagnostic Sciences,
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
\\
$\ast$ E-mail: acs4@cornell.edu, Phone: 607-254-1157, Fax: 607-255-4698
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

@@

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\textit{Streptococcus pyogenes} (SPY) has afflicted people around the globe
causing many diseases. Its closely related species, \textit{S.\ dysgalactiae}
subspecies \textit{equisimilis} (SDE), used to be known as veterinary pathogens
have been reported to share disease profiles with SPY.  While gene transfers
between the two species were shown in small portions of genomic loci, no study
on the gene transfer at genome-scale has been performed because of lack of
methodologies implemented or genomic data available. Recent advances in both
method development and genomic data allowed to study gene transfer between the
two species in genome scale.  Recipient bacteria of transferred genes can
substitute transferred genes for homologous parts of their genomes, or integrate
transferred genes into their genomes without replacement.  The former is called
recombining gene transfer, and the latter horizontal gene transfer. By
discriminating the two processes, We found that virulence genes were more
associated with horizontal gene transfer than recombining gene transfer.  We
also found a signficant signal of recombining gene transfer for the net
direction from SPY to SDE in the core genome.  Our study provides new insights
into gene transfer between SPY snd SDE in genome scale.

\section*{Introduction}

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}. However, it has been reported to cause serious and
life-threatening streptococcal diseases to humans traditionally associated with
GAS \citep{Brandt2009}.  Because of the overlap in disease profile and
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
well interact with each other by exchanging parts of their genomes.

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. Sachse et al.\ \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. Davies et al.\ \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. They also \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci. Ahmad et al.\ \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
Kalia et al.\ \citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal of
genomes available.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond such rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, Liu et al.\ \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of Karlin \citet{Karlin2001}. Luo et al.\ \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to Zhaxybayeva et al.\ \citet{Zhaxybayeva2006}. Hamady et al.\ \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
Caro-Quintero et al.\ \citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, Suzuki et al.\ \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

@@

Homologous recombining gene transfer could be studied more readily using the
core genomes of individual bacteria while horizontal gene transfer might be more
easily detected using the pan genomes. Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segments between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.  Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene duplication, loss, and
transfer events was developed by Doyon et al.\ \citet{Doyon2011}.  We noted that the first
approach could be used to infer transfer of genomic segments due to homologous
recombination, and the second could be useful in studying horizontal gene
transfer.

@@

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Recombination drived more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
model-based approach.  A genome alignment of the five genomes showed two
distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
SDD genomes were larger in length than those of SPY. Although the graphical view
of the genome alignment did not clearly show whether SDD was different from
either of SDE1 and SDE2 in the genome organization, a spcies tree inferred from
the genome alignment showed that the external SDD branch was attached to the
internal clade branch of SDE1 and SDE2 (Figure S\ref{fig:tree5}). The
distribution of the alignment block lengths was skewed to the right with mean of
4188 base pairs long (Figure S\ref{fig:blocksize}), which supported that the
model assumption of geometric block length distribution might be plausible in
\texttt{ClonalOrigin} method (see section \textbf{Materials and Methods} and
\citet{Didelot2010}).  The relative strength ($r/m$) of recombination compared
to mutations indicated that recombination-driven substitutions affected more
sites of the core genomes than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
species bacterial recombination process appeared to accelerate molecular
evolution of more sites than did point mutations process.

\subsection*{Recombination hotspots were observed in genomes}

@@

\subsection*{SPY and SDE were more closely related in about tenth of the core genome}

@@

\subsection*{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}

@@

\subsection*{Reduction of genes and divergence of SPY and SDE species}

@@

\subsection*{Virulence genes via recombining gene transfer}

@@

\subsection*{Association of virulence genes and horizontal gene transfer}

@@

\subsection*{Model-based approach or parsimony-based one for recombining gene
transfer study}

@@

\subsection*{UCSC genome browser tracks for bacterial gene transfer study}

@@

\subsection*{Units of recombining gene transfer}

@@

\subsection*{Transporters, transferases, aminoacyl tRNA synthetases, and
transposases}

@@

\subsection*{Simulations}

@@

\section*{Discussion}

@@

Were there unbalanced recombining gene transfer between the two species, what
were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
the species?  Kalia et al.\ \citet{Kalia2001} discussed restriction-modification system (RMS)
as a possible mechanism of the unbalanced gene flow.  Restriction-modification
system is a biological self-defense mechanism of bacteria, which is diverse in
prokaryote world.  Biological roles of RMS might be to maintain and control
species identity \citep{Jeltsch2003}.  Recently, Budroni et al.\ \citet{Budroni2011a} studied
that \textit{Neisseria meningitidis} phylogenetic clades were associated with
RMS, claiming that RMS modulated homologous recombination in the bacteria of
\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contained
restriction-modification type I system as well as type II system. The two
genomes of SDE contained only type II system. More tight self-defense system of
SPY may have made the species less capable of aquiring foreign genetic material
from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
may have been associated with different restriction-modification systems in the
two species \citep{Kalia2001}.

We reexamined the seven genes that Kalia et al.\ \citet{Kalia2001} had studied for net
directionality of gene transfer between SPY and SDE.  The internal fragments of
the six housekeeping genes (\textit{gki}, \textit{gtr}, \textit{murI},
\textit{mutS}, \textit{recP}, \textit{xpt}) were located in the core genome that
we investigated in this study.  Except for \textit{mutS} where Kalia et al.\ \citet{Kalia2001}
did not find recombination, our results were consistent with that of
Kalia et al.\ \citet{Kalia2001}.  Pinho et al.\ \citet{Pinho2010} reported recombination in the gene
\textit{parC}, in which we also observed moderate recombination.

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

\subsection*{Complete bacterial genomes}

Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
accessions: NC\_004070 \citep{Beres2002}, NC\_008024 \citep{Beres2006},
NC\_012891 \citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to
these samples as SPY1, SPY2, SDE1, and SDE2, respectively. The genome of
\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
Suzuki et al.\ \citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
ssp.\textit{\ equi} strain 4047 referred to as SEE \citep{Holden2009}.

\subsection*{Model-based method for recombining gene transfer}

We present a short summary of the statistical method developed by
Didelot et al.\ \citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
recombination. A detailed account was found in the two papers and in the user
guide available at \url{http://code.google.com/p/clonalorigin}.  We used
\texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with default
options to attempt to align the six genomes, which led to too small proportion
of core genomes. We had to remove the SEE sample from the genome alignment
although we kept it in the parsimony-based approach.  We then used
\texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
identified 276 blocks of sequence alignments with length threshold of 1500 base
pairs.  We excluded two of these blocks in a subsequent analysis because they
contained long consecutive gaps.  We applied \texttt{ClonalFrame} v1.1 available
at \url{http://www.xavierdidelot.xtreemhost.com} to the remaining 274 blocks,
and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
also estimates the fraction between recombination and mutation rates.
We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
and 10000 sampling generations to obtain the tree estimate with branch lengths,
and additionally seven independent MCMC runs to validate adequate convergence of
the first run.

Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
employed independently for each of the 274 alignment blocks as described in
Didelot et al.\ \citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
recombination rate, and recombinant tract length.  Two independent chains were
executed with $10^6$ burn-in generations and $10^7$ sampling generations. We
sub-sampled the result every $10^5$-th generations.  A few alignment blocks were
not successfully finished even after a month of computation.  Because those were
not expected to affect the global estimates of population parameters
significantly, we used only finished blocks to estimate mutation rate,
recombination rate, and recombinant tract length.  The global estimators of the
three population parameters were medians weighted over block lengths, which
down-weighed less reliable estimates obtained from shorter alignment blocks.
One of the two independent chains was used to obtain the block-wise mean
estimates, and the other chain was used to validate convergence of the first
stage of MCMC.  Two independent MCMC runs resulted in similar estimates of the
three population parameters (Table S\ref{tab:three}).  The second stage of MCMC
with \texttt{ClonalOrigin} was executed with the three population parameters
fixed to the estimated values in the first stage.  Two independent MCMC chains
were also executed with a total generation being $1.1\times10^8$. The first
$10^7$ generations of each chain were discarded as burn-in. The remaining $10^8$
generations were sub-sampled every $10^5$-th generation.

\subsection*{Parsimony-based method for horizontal gene transfer}

While \texttt{ClonalOrigin} could be used to understand recombining gene
transfers within the conserved core genome, we sought to understand the history
of the remainder of the genome where more complex evolutionary processes occur,
such as gene duplications, losses, and horizontal transfers.  In recent years,
several methods have been developed for inferring these evolutionary events in
gene families by reconciling gene trees with a species tree
\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we complemented our
study of recombination by using \texttt{Mowgli} \citep{Doyon2011} to
characterize events that can change the copy number of a gene, such as
duplications, losses, and non-recombining horizontal gene transfers.  In
contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} is parsimony-based for
inferring events and uses pre-determined gene clusters where genes are treated
as atomic units.  The advantage of using \texttt{Mowgli} was that it did not
assume the sequence data contains a single copy of each homologous segment, and
it attempted to explain different copy numbers by a series of gain and loss
events.  We obtained gene family clusters defined in Suzuki et al.\ \citet{Suzuki2011} using
the \texttt{OrthoMCL} method \citep{Li2003} over the full set of the six genomes
including \textit{S.\ equi} subsp.\textit{\ equi} strain 4047 (SEE).  For each
family cluster, the corresponding protein sequences were aligned using the
\texttt{MUSCLE} program \citep{Edgar2004a} and nucleotide alignments were
constructed by mapping nucleotide sequences onto each protein sequence
alignment. We then identified recombination within gene alignments using the
Single Breakpoint Recombination (\texttt{SBP}) method from the \texttt{HyPhy}
package \citep{KosakovskyPond2006}.  Using the Akaike Information Criterion
(AIC), \texttt{SBP} inferred that 47.5\% (1,094) of gene families contained at
least one topology-changing recombination. Each of these families was then split
into two gene fragments.  Gene trees were constructed from nucleotide alignments
for each gene fragment using the \texttt{RAxML} program \citep{Stamatakis2006},
and were rooted by minimizing inferred duplications and losses.  Each rooted
gene tree was then reconciled using the \texttt{Mowgli} program
\citep{Doyon2011}, which inferred the most parsimonious gene duplication, loss,
and transfer events for each family.  In the case of multiple maximum
parsimonious reconciliations, one was chosen uniformly at random.  Unless
specified otherwise, default setings were used for all the programs.  Since the
method of \texttt{Mowgli} did not directly model recombination events,
recombination-based transfers were inferred in a post-processing step, by
checking whether the transfer was coupled with a loss of a homologous gene in
the recipient branch (indicating gene replacement rather than gain).  A transfer
event was classified as non-recombining if the destination species of that
transfer had an extant descendant that contained copies of that genes which were
descendants of the transferred gene as well as copies which were
non-descendants.  We noted that this parsimonious approach was imperfect and
would tend to undercount events, especially recombining gene transfer events,
which were modeled using two events: a transfer and a loss.  Nevertheless, this
analysis allowed us to study non-recombining gene transfers, which were not
modeled by \texttt{ClonalOrigin}, and to validate some of the
\texttt{ClonalOrigin} results on recombining gene transfer.

\subsection*{Functional categories association}
